Supplement\: Considering scale in multilayer animal social networks

### 2020-06-14 paper/supplement.Rmd
fst supplement

---
title: Considering scale in multilayer animal social networks
author: Alec L. Robitaille^1\*^, Quinn M.R. Webber^2^, Julie W. Turner^1^, Eric Vander Wal^1,2^
date: |
  | ^1^Department of Biology, Memorial University of Newfoundland, Canada
  | ^2^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland 
  | \*Corresponding author: Alec L. Robitaille (robit.alec@gmail.com)
output:
   bookdown::github_document2: default
   # bookdown::word_document2:
   #   reference_docx: template.docx
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
bibliography: ../references/knitcite.bib
csl: csl.csl
---

@@

# Abstract
Scale remains a foundational concept in ecology. Spatial scale, for instance,
has become a central consideration in the way we understand landscape ecology
and animal space use. Meanwhile, social processes can also scale from fine-scale
interactions to co-occurrence to overlapping home ranges. <!-- TODO: add
temporal scale --> Multilayer networks promise the integration of monolayer
animal social networks with the complexity and importance of animal space use
and movement in heterogeneous landscapes. Despite the complex interplay between
social networks and how animals use space, there remains an important biological
and methodological gap in our understanding of an animal’s perception of scale.
Here we discuss the role of scale in the context of multilayer networks and
provide a social ungulate case study to illustrate the role of social, spatial,
and temporal scale on multilayer processes. We also present perspectives on
future development and applications of scale in multilayer networks with respect
to: phenotypes and individual fitness, movement ecology and collective movement,
and habitat selection and space use. Effective integration of social and spatial
processes, including biologically meaningful scales, within the context of
animal social networks is an emerging area of research, as such our contribution
incorporates perspectives on how the social environment and spatial processes
are linked across scales in a multilayer context.

## Keywords
Social network analysis, Landscape ecology, Movement ecology, Space use

# Introduction
Our inference regarding the processes underlying ecological patterns
are directly linked to the scale at which they are observed [@Levin_1992;
@Allen_2015]. Animals are influenced across scales, through processes such as
species-specific as trophic interactions and resource availability
[@Legendre_1993; @Chave_2013] and global climate and productivity [@Field_2009].
Multiscale ecology has been integrated into species distribution modelling
[@Elith_2009], habitat selection [@Mayor_2007], and food webs [@Sugihara_1989],
among others. Meanwhile, in the context of animal behaviour, certain behaviours
are scale-dependent; for example, acoustic communication in birds varies with
spatial scales [@Luther_2009]. We posit that it is important to consider that
animal behaviour within the social environment will be scale-dependent.

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods [e.g. @Carter_2015], whereas social association
represents shared space use by members of the same social group [Figure
\@ref(fig:spacetime), @Franks_2009]. Further, for social interactions or
associations to occur individuals must share space, and thus have overlapping
home ranges [@Vander_Wal_2013]. In the context of the social environment, home
range overlap is an example of how animals share space over coarser spatial and
temporal scales (e.g. Piza-Roca et al. 2018). Although the social environment
clearly scales spatially and temporally, it remains unclear whether coarser
scales of sociality, including social association and home range overlap, vary
predictably with social interaction [@Castles_2014;
@Farine_2015a]. Thus, we define social scale as the type of social relationship
observed that can be explicitly defined and measured [@Farine_2015a].

Animal social network analysis is a well-developed tool used to measure the
relationships of individuals and organization of social systems [@Krause_2009;
@Wey_2008; @Croft_2011]. Social network analysis provides insight into the
structure of social communities and social network properties, which can
influence population dynamics and evolutionary processes [@Pinter_Wollman_2013;
@Kurvers_2014]. Despite the widespread use and innovation of traditional social
network analysis [@Webber_2019], it typically considers a single scale of
sociality in a given network (monolayer), drastically simplifying the
complexity of animal social systems [@Finn_2019].

Unlike traditional monolayer social networks, multilayer networks explicitly
consider social systems across scale-dependent contexts
[@Pilosof_2017]. Multilayer networks are made up of multiple layers, each
representing (i) different classes of individuals, e.g. male or female, (ii)
types of behaviours, e.g. grooming, travelling, or foraging, (iii) spatial
areas, e.g. local or regional, (iv) or temporal windows, e.g. daily or seasonal
[@Kivela_2014; @Porter_2018]. Multilayer networks are relatively novel to
studies of animal behaviour [@Silk_2018; @Finn_2019], although they have been
used to describe multidimensional human social systems, complex transportation
networks, and neural networks [@Silk_2018]. The role of scale within multilayer
social networks remains unexplored but provides a unique opportunity to develop
novel understanding of variation in the social, spatial or temporal scales associated
with social systems.

Our examinations is motivated by behavioural, landscape, and spatial ecology to
build a conceptual and analytical framework for scale-dependent multilayer
networks, and we apply this framework to a case study of caribou (*Rangifer
tarandus*) in Newfoundland, Canada [@Peignier_2019]. We use this case study to
address two principal questions associated with scale-dependent multilayer
networks. First, to what degree does the social scale of relationships, spatial
scale of landscapes, and temporal scale of analysis influence our ability to
interpret complex social systems? Second, can variation in the social, spatial,
and temporal scale improve predictions associated with biological differences in
seasonal resource availability and social association? We conclude with
perspectives for further development and application of scale in multilayer
networks with respect to phenotypes and individual fitness,
movement ecology and collective movement, and habitat selection
and space use.
<!-- TODO: we answer the first question in the discussion, but not the second-->
<!-- TODO: Webber reference should be 2018, but defaulting to 2017-->

(ref:spacetime) Space-time diagram representing variation the relative spatial and temporal extent required for different types of social and communication processes for four species, including spotted hyaena, passerine birds, sleepy lizards, and elephants. Spatial and temporal extent for social interactions, e.g. mating, grooming, or aggression, are similar for most species because physical contact between two individuals is required for many social interactions. The logical extension is that spatial and temporal extent for social interaction is hierarchically nested within the spatial and temporal extent for social association because individuals must share space to interact. By contrast, different species have potential for greater spatial and temporal extents, for example, temporal extent for vocal communication is similar for most species because most vocal calls only persist in the environment for seconds, but spatial extent for vocal communication is highly variable with elephant calls extending the great distance and passerine calls extending the shortest distance.

<!-- TODO: units of spatial scale? --> 
<!-- TODO: citations? -->

# Methods

## Caribou socioecology
<!--Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on predicted social (visual and auditory perception), spatial
(habitat type), and temporal (seasonal) scales.--> Caribou are gregarious ungulates
with fission-fusion dynamics [@Lesmerises_2018] that display temporal and
spatial variation in social networks [@Peignier_2019]. In winter, caribou dig
holes in the snow, termed craters, to access forage [@Bergerud_1974]. Cratering
is presumed to be a costly behaviour and as a result, caribou tend to occupy and
re-use craters once they are established. <!-- and there is considerably less access
to forage than when the landscape is snow-free.--> Importantly, craters exist on
the landscape at multiple scales: the crater scale (a single crater), the
feeding area scale (multiple craters in close proximity), and the winter range
scale (all craters within an individual’s range) [@Mayor_2009]. Access to forage is 
seasonally heterogeneous depending on snow cover and, in winter, variable distribution of craters on the 
landscape. <!--We partitioned data into two
discrete seasons: summer (3 August - 20 September 2017) and winter (1 January – 17
February 2018) based on caribou socioecology [@Peignier_2019].-->

## Caribou location data
We used GPS location data collected from Fogo Island caribou in between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. Adult female caribou were immobilized and fitted with global positioning
system (GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M
collars, 1,250 g) as described by [@Schaefer_2013]. Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
to remove all erroneous and outlier GPS fixes following
[@Bjorneraas_2010]. We did not collar all caribou in the herd; however, the
proportion of marked adult females was ~10% of all adult females and ~5% of all
individuals in the herd. We assumed these individuals were randomly distributed
throughout the population.

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification
included nine habitat types at 30 m resolution. We reclassified the land cover
types into three categories using the R package `raster` [@Hijmans_2019]:
foraging, open, and forest habitats. Foraging habitat consisted of lichen, open
habitat consisted of wetland, rocky barrens, and anthropogenic, while forest
habitat consisted of coniferous forest, conifer scrub, broad leaf forest, and
mixed-wood forest. Water habitat was excluded from all subsequent analyses.

## Caribou multilayer social networks

### Network layer construction
<!--We generated a multilayer social network consisting `r DT[, uniqueN(get(idcol))]`
individual nodes and two aspects (season and land cover). Two season
elementary-layers (summer and winter) and three land cover elementary-layers
(forage, forest, and open) combined for a total of six layers. Single layers were
composed of individuals linked by social associations, for each season and land cover combination.-->

We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r version$version.string` [@R_Core_Team_2019]. We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the spatial distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a spatial distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018].

<!--In
all networks, individual caribou were represented as nodes, associations between
individuals in a given habitat type and season were represented as intralayer
edges, and connections between the same individuals across seasons and within
habitat classes were represented as interlayer edges (Figure \@ref(fig:ml)) .-->
<!-- eg this is connecting contexts used to calculate layer similarity -->

We weighted edges of social networks by the strength of association between
caribou using the simple ratio index [SRI, @Cairns_1987]:

$$ SRI = \frac{x}{x + y_{AB} + y_{A} + y_{B}} $$

where x is the number of fixes where individuals A and B were in the same group,
$y_{A}$ is the number of fixes from individual A when individual B did not have
a simultaneous fix, $y_{B}$ is the number of fixes from individual B when
individual A did not have a simultaneous fix, and $y_{AB}$ is the number of
simultaneous fixes from individuals A and B that were separated by more than the
spatial distance threshold [@Farine_2015].

### Network metrics 
<!-- TODO: expand + simplify + examples individual A, B etc -->

We used a series of metrics to characterize the multilayer networks, focusing on
the role of individuals within and across layers, and the similarity of
different layers. To measure the role of individuals within layers, we
calculated degree centrality and graph strength. Degree centrality is the number
of direct connections an individual has to other individuals in a network layer.
Graph strength is the degree weighted by the strength of association, in this
case SRI, in each layer. Extending degree centrality, we calculated multidegree,
the sum of degree centrality of individuals across layers  [@Berlingerio_2012;
@Kivela_2014]. <!--We evaluated the relevance and redundancy of layers by
combining the multidegree and neighbourhood to calculate layer relevance and
connective redundancy. Layer relevance is the proportion of neighbours present
in each layer compared to the neighbourhood for each individual and connective
redundancy is the proportion of neighbours repeated in multiple layers, and
[@Berlingerio_2012].--> Finally, we measured the similarity of layers by
calculating edge overlap. Edge overlap is defined as the proportion of edges
present in each network layer out of all observed edges  [@Battiston_2014].
<!--between seasons, across land cover classes (e.g. summer open compared to
winter open) using the Pearson correlation coefficient [@Brodka_2018].-->
Network metrics were calculated using the R packages `igraph` [@Csardi_2006],
`asnipe` [@Farine_2019], `spatsoc` [@Robitaille_2019] and `data.table`
[@Dowle_2019].

## Varying scale in multilayer networks
<!--

social scale: variable is spatial distance threshold
spatial scale: variable is land cover resolution
temporal scale: variable is temporal window
temporal scale: variable is number of observations
-->

### Social scale

We generated multilayer networks across a series of spatial distance thresholds
for group assignment. Spatial distance thresholds (5, 25, 50, 75, 100, 250 and
500 m) represented a range of visual (short distance) to auditory (long
distance) sensory modalities of caribou. Multilayer networks consisted of the
social association between `r DT[, uniqueN(get(idcol))]` individuals across
three habitat layers (open, forest and forage) for the entire study period. At
the finest scale, individuals within 5 m of one another were considered in the
same group, whereas at the coarsest scale, individuals within 500 m of one
another were considered in the same group. We calculated graph strength and
multidegree for each habitat layer and spatial distance threshold, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial distance
threshold.

### Spatial scale
<!-- TODO: Increased spatial scale of land cover corresponds to xyz -->
To assess the influence of spatial scale, we aggregated the land cover raster using the R package `grainchanger` [@Graham_2019] across a series
of scales (100 - 1000 m by steps of 100 m). The land cover raster was aggregated
using a modal moving window method using a circular window corresponding to the
above scales [@Graham_2019a]. We assume 30 m and 100 m represents fine-scale
decision making for caribou during foraging, while re-sampling at 500 m , 750 m
and 1000 m represents the scale at which caribou tend to select and avoid
habitat (Bastille-Rousseau et al. 2017). For each spatial scale, we generated a
multilayer network consisting of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (open, forest and
forage) for the entire study period. Social associations of individuals were
defined by spatiotemporal overlap within 5 minutes and 50 m. We calculated graph
strength and multidegree for each habitat layer and spatial scale, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial scale.

### Temporal scale
#### Time windows

We used a temporal multilayer network to assess the seasonality of caribou
sociality. Multilayer networks consisted of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across 20 ordinal sample periods of 35-36
days taken from the entire study period. Social associations were defined by
spatiotemporal overlap of individuals within 5 minutes and 50 m. In each time
window, we calculated individual graph strength and multidegree, and across the
series of networks, we calculated edge overlap to determine proportion of total
edges observed in each time window.

#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies. To investigate the influence of number of observations, we generated multilayer networks of social association between `r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (open, forest and forage) using $N$ randomly selected observations. We first generated a maximum of 1000 timesteps and iteratively used $N$ of these ($N$ = 10 - 1000 observations by steps of 10), to ensure previously included timesteps, and resulting associations, were accumulated with subsequent observations to mimic collection of observational data. Within each multilayer network with $N$ observations, we calculated individual graph strength and multidegree. We calculated
edge overlap across the entire series of networks to determine proportion of
total edges as number of observations increased.

# Results
<!-- TODO: edge overlap can indicate if some associations are dependent on context -->
<!-- TODO: Add intro/general to results?-->
<!-- major points then details -->

Individuals in multilayer networks became more connected as the
spatial distance threshold increased. Mean
graph strength differed across habitat layers with similar and higher graph
strength in open and forage layers, and lower strength in forest layers (Figure
\@ref(fig:socres)). Individual graph strength increased sharply between 5 m and
100 m in all three habitat classes followed by a decrease in rate after 100 m. These
results suggest the optimal social scale at which groups should be assigned is
likely somewhere between ~20 - 100 m based on the rate at which network metrics
stabilized. For ungulates, groups are often assigned based on 50 m threshold
(Lingle 2003; Lesmeries et al. 2018; Peignier et al. 2019), suggesting that
given similar quantities of data, ~50 m is likely an appropriate social scale
for generating social networks in caribou and similar species.

Increasing land cover resolution resulted in a decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats at higher resolution (Figure \@ref(fig:lcres)). Graph strength
and edge overlap was relatively consistent across land cover resolutions for
forest and open layers (Figure \@ref(fig:lcres) D, E). Edge overlap decreased
sharply between 30 m and 600 m in forage layers, afterwards remaining stable to
1000 m (Figure \@ref(fig:lcres) D). Graph strength was more variable within
forage layers across land cover resolutions than open and forest (Figure
\@ref(fig:lcres) D). The proportion of relocations in forest and open habitats
increased with increasing spatial resolution, while decreasing in forage habitat
(Supp. 1 <!-- todo-->).  Overall, these results indicate the importance of
matching land cover resolution to scale of selection as well as ensuring the
land cover product used has a resolution fine enough to detect landscape
features and habitats relevant to study species, such as important patchy and
rare habitats.

<!-- TODO add? each ordinal network layer consisted of avg 380 fixes * ID -->

The ordinal multilayer network generated from discrete time windows indicate
within-year periods of increased caribou sociality. Edge overlap was higher in time windows 1, 6-11 and 15-20 than time windows 2-5 and 12-15  (Figure \@ref(fig:temp) A). These periods between approximately October and May

increasing graph strength with overall peak in 2018 but still increases in 2019
note graph connectedness is evident panel B

<!-- Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time window 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.  -->
These
results suggest that an increasing window length, especially for the summer
season, may be resulting in less precisely defined seasons that are not clearly
capturing periods of aggregation and dissaggregation related to resource
availability.

changing the position of the
time window in this case results in a mismatch between expected periods of
resource seasonality and corresponding sociality, particularly for the summer,
highlighting the importance of carefully selecting temporal scale of analysis.

<!-- TODO: reduce to this version -->
As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics appeared to stabilize. Mean graph
strength was consistently weak in summer layers but varied in winter layers at
low number of observations, becoming stable at ~75 observations (Figure
\@ref(fig:nobs)a). Layer relevance in the summer was consistently low, but
winter layers increased between approximately 10  and 100  observations,
remaining stable with additional observations (Figure \@ref(fig:nobs)b). Like
coarse land cover resolution (Figure \@ref(fig:lcres)), low numbers of
observations were not sufficient for capturing individuals in all contexts and
therefore layer similarity could not be calculated below ~30 observations for
forest, approximately ~80 observations for open (Figure \@ref(fig:nobs)c). Layer
similarity was not possible to calculate for forage layers because there were no
connected individuals in summer forage. Once sufficient observations were
obtained, layer similarity showed some variability with increasing numbers of
observations for forest and was relatively stable in open layers. Connective
redundancy increased dramatically between ~10 and ~50 observations before
stabilizing with additional observations included (Figure \@ref(fig:nobs)d).
Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

<!-- TODO: I'm not sure this is relevant since this is the main way it
is already done, can we make this thought more relevant to picking scales or
multilayer networks in general? I think it's important for building robust
networks -->

<!-- Figure: metrics by social scale -->
(ref:socres) Varying scale in spatial distance threshold. For each spatial distance threshold (100 - 1000 m by steps of 100 m), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). Graph strength (mean of individuals in bold) showed a sharp increase between 5 m and 100 m for all habitat classes. After 100 m, graph strength continued to increase at a slower rate. Open and forage layers showed higher variation in graph strength between individuals while the forest layer had less variation and lower mean graph strength.

<!-- Figure: metrics by land cover resolution -->
(ref:lcres) Varying spatial scale of land cover raster. For each land cover resolution (5, 25, 50, 75, 100, 250 and 500 m), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A, B, C) Three land cover rasters: the original resolution (30 m) and two aggregated rasters (500 m and 1000 m). D) Edge overlap of habitat layers across spatial resolutions. Open and forest layers show consistent edge overlap with increasing spatial resolution while forage shows a sharp decline in edge overlap between 30 m and 600 m. Above 600 m, the forage layers are stable. E) Graph strength (mean of individuals in bold) within habitat layers across spatial resolution. Forage layers showed high variation in graph strength with many individuals dropping to 0 after 500 m and others increasing with spatial resolution. Open layers and forest layers were relatively stable across spatial resolutions, with higher mean graph strength in open compared to forest.

<!-- Figure: metrics by number of observations -->
<!-- TODO: update to temp -->
(ref:nobs) Varying scale in number of observations. Multilayer networks were generated using a  sample number of observations (10 - 1000 observations by steps of 10) of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A) Graph strength was highly variable at low number of observations (less than 100) for all habitat layers and relatively stable after 200 observations. B) Edge overlap increased for all habitat layers with increasing number of observations. Open and forage layers had higher edge overlap than forest layers. C) Multidegree increased for all individuals with increasing number of observations. Individuals showed high variability (`r nobs[nobs == max(nobs), paste(range(multideg), collapse = ' - ')]`) at the maximal number of observations (1000).

<!-- Figure: Time window -->
(ref:temp) Varying temporal scale in multilayer networks. Multilayer networks were generated of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland within ordinal time windows (20 ordinal sample periods of 35-36 days) between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. A) Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time window 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.

<!-- TODO: a bit more for g strength? -->

# Discussion

<!-- TODO: caveat social scale could also be diff types of relationships but GPS but point to hmm -->

The integration of behavioural and ecological theory has shown the value of
considering the importance of various ecological disciplines to better
understand and predict animal social networks, including demography
[@Shizuka_2019], evolutionary and quantitative genetics [@Fisher_2017], disease
ecology [@Silk_2017], and animal personality [@Sih_2018]. While the number of
studies integrating social network theory into a broader array of behavioural
and ecological sub-disciplines is impressive, only recently has the role of
animal space use been considered in describing the causes and consequences of
the social environment in the context of (1) phenotypes and individual fitness
[@Webber_2018], (2) movement ecology and collective movement [@Bode_2011], and
(3) habitat selection and space use [@He_2019]. Importantly, much of the work
linking space use and the social environment has been theoretical and empirical
studies are lacking [but see @Peignier_2019; @Strandburg_Peshkin_2015;
@Strandburg_Peshkin_2017]. Here, we discuss the role of scale and multilayer
networks in the emerging integration of social and spatial ecology within the
behavioural ecology literature.

## Phenotypes and Individual fitness

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Multilayer networks
provide a way to bridge these different scales so that we can examine social
phenotypes in more holistic ways rather than looking at each scale separately as
has traditionally been done. From a monolayer network perspective, social
phenotypes are increasingly linked to important fitness for individuals of
various species, including primates [@Brent_2017; @Thompson_2019], dolphins
[@Stanton_2012; @Stanton_2011], and birds [@McDonald_2007; @Royle_2012]. Both
social scales, types of interactions, and temporal, seasons and stages of life,
have emerged as important for understanding fitness consequences in the context
of social network positions [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011].

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues unlike in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

Temporal scale is also important for understanding the repercussions of social
phenotypes. Varying the time window length and position for caribou, we found
there was a lot of variation in the multilayer network metrics (Figure
\@ref(fig:winlen), Figure \@ref(fig:winpos)). These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest. For instance, social phenotypes can change over ontogeny
in ways that can affect fitness [@Berger_2015; @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can make defining appropriate temporal
scales for testing the fitness implications of an individual’s social phenotype
over different stages of ontogeny even more difficult when they vary by
individual. However, considering fine temporal scales provide insight across
developmental stages that may otherwise be missed [@Turner_2017]. Defining
temporal scales is a challenge in multilayer networks just as in monolayer
networks; however, multilayer networks provide metrics that enable researchers
to look at similarities and differences in separate layers representing
different stages of ontogeny, or seasons as in our case study, which can be
difficult to detect.

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful reduction of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a framework
to test hypotheses linking social phenotypes and fitness across temporal scales.
Because of the non-independent nature of social interactions and associations,
especially over time, it is hard to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

Conversely, picking the proper temporal scale is important but a challenge
because all layers in the multilayer network should reflect the same time
period, and the period should be long enough for each layer to have enough data
for robust networks [@Farine_2017]. These considerations may be limiting factors
when answering questions that require time frames that are relevant to different
individuals, for instance. In both mono- and multilayer networks, the effect of
scale on the relationship between an individual’s social network position and
fitness need to be considered based on the hypotheses being tested and
logistical constraints.

## Movement ecology and collective movement

Social processes, movement decisions, and space use are inextricably linked and
multilayer networks represent a potential analytical framework for linking these
processes [@Mourier_2019]. Movement as a collective group can reduce per capita
predation risk via detection-dilution trade-offs and improve information
transfer about the quality or location of resources to other group members
[@Jolles_2019]. In our case study, we integrate proximity based social networks
that inherently rely on animal collective movement, and, in general, we found
that movement is likely an important mechanism driving social network
connectivity within and between habitat types.

Although multilayer animal social networks are relatively novel [@Silk_2018],
movement behaviour within multilayer networks has been considered. Spatial
networks are physical locations (nodes) which are connected in a network when
individuals move between said locations (edges). For example, spatial networks
were applied to monitor movement of sharks (*Carcharhinus amblyrhynchos* and
*Carcharhinus melanopterus*) between fixed location autonomous arrays. In this
example, nodes represent locations within the array, *intralayer edges*
represent individual movement frequency, and *interlayer edges* represent
contact probability among individuals at a given node [@Mourier_2019]. Although
movement between fixed arrays is inherent, networks are not constructed based on
continuous measures of dyadic or collective movement [@Long_2014]. Our case
study builds on the work of Mourier et al. (2019) and explicitly integrates
continuous and simultaneous measures of animal movement through space and time
as layers of a multilayer social network.

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine
gunieafowl (*Acryllium vulturinum*), GPS movement data were used to generate
monolayer association networks within a multilayer network [@Papageorgiou_2019].
However, despite potential to generate networks in different spatial or social
contexts, layers in the multilayer network represented social associations
within and between groups. Continuous collection of animal relocation data using
GPS technology can be used to measure how groups of animals make collective
movement decisions in the context of their local environment
[@Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017] and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups [@Bode_2011]. This movement layer could be made at a
fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale [global migration:
@Flack_2018]. Our case study integrates variation in temporal and spatial scale
to assess the role of scale in predicting multilayer networks. As one might
expect, we found increasing social threshold results in saturation of potential
social connections (Figure \@ref(fig:socres)).<!--TODO: what??, while increasing temporal scale had little effect
on habitat-specific networks (Figure \@ref(fig:))-->. Another potential application would be
to generate social networks during times when animals are engaged in different
behaviours, including travelling, interacting, or foraging [@Muller_2018;
@Finn_2019] and construct a multilayer network based on behaviour-specific
monolayer networks [e.g. @Smith_Aguilar_2019].

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology has
advanced to the point where relocations can be programmed to occur as frequently
as each second [@B_rger_2016], a practice which has been applied to questions of
collective movement [e.g @Strandburg_Peshkin_2015]. Meanwhile, remote sensing
technology, including unmanned aerial vehicles and satellites, can be used to
track the fine-scale movements of groups over large distances and time scales
[@Torney_2018]. Using technology to infer behaviour has therefore become
standard operating procedure for many species in many systems. Our ability to
collect and process remotely sensed data and infer social processes based on
animal movement provides an opportunity to use multilayer social network
analysis to answer complex questions about animal socioecology.

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that explicitly
quantify movement parameters could be integrated. For example, edges could
represent measures of dyadic or collective movement based on dynamic interaction
metrics [@Long_2014]. Integrating movement layers with habitat-specific social
association layers could further illuminate the role of movement as a driver
of the social and spatial environments an individual experiences [@Webber_2018].
In our case study, we generated multilayer networks based on habitat-specific
monolayers. Coordinated movement of dyads, or groups, could vary based on
habitat, and the approach we outline could be used to explore mechanisms linking
the social and spatial environments. For example, social processes may be an
emergent property of the landscape because animals aggregate at resources, such
as waterholes [@Chamaill_Jammes_2008]. Movement to, from, or within habitat
patches or territories can also contribute to the formation of the social
environment [@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*)
males interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a multilayer
network framework is a logical next step for a burgeoning field. In particular,
when association networks are habitat-specific, we suggest the integration of
movement and space use in multilayer networks provides novel insights into the
effects of habitat configuration on the formation of the social environment
[@He_2019; @Webber_2018].

## Habitat selection and space use
Social structure and habitat selection are intimately linked: individuals must
share space to interact and the distribution of resources influences group
dynamics and home range size [@Webber_2018]. Social structure is influenced by
resource availability and seasonality [@Peignier_2019], configuration [@He_2019]
and fragmentation [@Banks_2007]. Habitat selection, the disproportionate use of
available resources or landscape features by animals [@Manly_1993; @Boyce_1999],
is a scale-dependent process [@Mayor_2009]. Multilayer networks represent a
novel framework for considering the influence of habitat selection and space use
on sociality across scales.

>>>>

---
title: Supplement\: Considering scale in multilayer animal social networks
author: Alec L. Robitaille^1\*^, Quinn M.R. Webber^2^, Julie W. Turner^1^, Eric Vander Wal^1,2^
date: |
  | ^1^Department of Biology, Memorial University of Newfoundland, Canada
  | ^2^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland 
  | \*Corresponding author: Alec L. Robitaille (robit.alec@gmail.com)
output:
   bookdown::github_document2: default
   # bookdown::word_document2:
   #   reference_docx: template.docx
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
bibliography: ../references/knitcite.bib
csl: csl.csl
---

@@

Table S1: Time cut definitions

```{r}
knitr::kable(timedefs)
```

Figure S1: n individuals

```{r}
knitr::include_graphics(')
```
